Work done during my PhD would not have been possible without the contribution of many people. This dissertation is indeed a baby which required a village to take care of.

I would first like to thank my advisors who saw this dissertation through. None of the work presented here would have been possible without Andrew Cohen and Jeffrey Starns who over countless meetings have helped me hone my projects, and ensured scientific rigor. They also let me work on whatever I wanted, provided the resources I needed, and generally made research fun. I would also like to thank Youngbin Kwak and Meghan Huber who served on my dissertation committee -- first for continuing on it as I changed projects and for providing invaluable feedback that significantly improved my dissertation. My first advisors in grad school at UMass; David Huber and Rosie Cowell taught me the fundamental steps of research in my early years as a PhD student really honed my ideas and allowed me to grow as a researcher. Other cognitive faculty at UMass Amherst provided invaluable experience in academic presentation through brown-bag questions and through courses I took over the years. 

I would next like to thank my \st{collegues} friends in the cognitive program throughout the years: Sean, Anna, Melisa, Mar, John, Jerome, Michael, Yun, Chiungyu, Kuan-Jung, and Sandarsh for being a stable presence in this intellectual journey. You made this program feel like home. I would also like to call out Natasha and Kuan Jung, my co-hort mates who went through the roller coaster with me. 

Weekly board games in PBS is one of the most consistent social activity I developed is at UMass and the group of friends have practically served as my family away from home. Anna, Clara, Tori, Sean, Trina, Fran, Ramiro, and Kuan Jung; your dedication to put up with my, ahem, eccentric, hobby has never failed to lift my spirits and always provided with an outlet I look forward to every week. The last two years of my grad school would not have been as much fun if not for you and I am glad that this tradition is set to continue going forward. Finally, I'd be remiss if I did not mention the `tiny living room' folks; Ramiro, Sandarsh, and Hyejoo for being my first non-work friends at work. 

Several people helped me through grad school before and through the pandemic -- Aarohi, Meet, Rik, Sohini, Pracheta, and Princy, I will never forget the absolute random conversations, and trips that provided a venue to turn my mind off work and let me stay sane. Kunal, while you were not at UMass, you have never let go of this friendship with regular check ins, and truly showed me the value of real friendship. I don't know how you do it, but please continue doing it.

Manasa -- I can write another dissertation to describe your tireless support. Instead, I will just say that absolutely no part of the dissertation would have been possible without you. You helped me in literally everything I wanted your help. From non-work ups and downs, to serving as an initial ideas sounding board, to helping me with analyses and programming, to feedback on my writing, and through celebrating me, you have seen it all, done it all and been my consistent cheerleader in these six years. I could not have chosen a better human being to do so. Thank you.

Finally, I would not have been here if not for continued, unwavering support from my family. Mummy, pappa, diku, Nishita, and Aaru -- thank you for letting me do something regardless of whether it is clear what it is I exactly do. Your faith in me has been the key to make this work happen.

